%<meta name="verify-v1" content="7dc0e9u+pSARcoW2BNthVV9GdlvC25zbT4CU8iFkNZo=" />
%latex2html -split 0  -no_navigation -no_subdir -info 0 CV.tex
% General template for CV
% Angel Lopez-Urrutia Lorene
% ----------------------------------------------------------------------

\documentclass[a4paper,12pt]{article}
\usepackage[latin1,latin9]{inputenc} % spanish punctuation
\usepackage{times}
%%  \usepackage{graphicx}
%% \usepackage[sort&compress]{natbib}
\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%\raggedbottom
\raggedright

\newcommand{\lineunder}{\vspace*{-8pt} \\ \hspace*{-18pt} \hrulefill \\}
\newcommand{\header}[1]{{\vspace*{14pt}\hspace*{-10pt}\vspace*{6pt} \textbf{\Large{#1}}} \vspace*{-6pt} \lineunder \vspace*{6pt}}
\newcommand{\header}[1]{{\hspace*{-10pt}\textbf{\Large{#1}}} \vspace*{-2pt} \vspace*{2pt}}

\newcommand{\contact}[3]{
\vspace*{-8pt}
\begin{center}
{\LARGE {#1}}\\
#2 \lineunder 
#3
\end{center}
}

\newenvironment{achievements}{\begin{list}{$\bullet$}{\topsep 0pt \itemsep -2pt}}{\vspace*{4pt}\end{list}}
\newcommand{\schoolwithcourses}[4]{
 \textbf{#1} #2 $\bullet$ #3\\ 
#4 $\bullet$  Selected Coursework:\\
\vspace*{5pt}
}
\newcommand{\school}[4]{
 \textbf{#1} #2 $\bullet$ #3\\ 
#4 \\
}
% END RESUME DEFINITIONS




\begin{document}
\small
\smallskip
\vspace*{-44pt}

\contact{\'{A}ngel L\'{o}pez-Urrutia}
{Centro Oceanogr\'{a}fico de Gij\'{o}n, Instituto Espa\~{n}ol de Oceanograf\'{i}a\\
 Avda. Pr\'{i}ncipe de Asturias 70 bis, Gij\'{o}n, Asturias, E-33212, Spain\\
Tel: +34 985 308 672 -- Fax: + 34 985 326 277 -- alop@gi.ieo.es}

\header{Education}
\underline{European Ph.D.} \textit{(With Distinction)} October 2002.\\
\underline{Ph.D. by the University of Oviedo} (Spain; Ecology program) October 2002.\\
Title: \textit{In situ evaluation of the ingestion rates of the appendicularian community: trophic impact and ecophysiological implications}. Advisor: J.L. Acu\~{n}a.\\
\underline{B.Sc. in Biology} (Zoology program) July 1997. University of Oviedo.\\

\header{Profesional Employment}
\underline{Senior Scientist}. Instituto Español de Oceanografía. Centro Oceanográfico de Gijón. (September 2007-).\\
\underline{Ramon y Cajal Fellow}. Instituto Español de Oceanografía. Centro Oceanográfico de Gijón. (December 2006-September 2007).\\
\underline{PI of Spanish Ministry of Science project}. Instituto Español de Oceanografía. Centro Oceanográfico de Gijón. (January 2005-December 2006).\\
\underline{Principado de Asturias fellow}. Instituto Español de Oceanografía. Centro Oceanográfico de Gijón. (July 2003-January 2005).\\
\underline{Zooplankton ecologist}.  U.K. Natural Environmental Research Council - Plymouth Marine Laboratory. (October 2000-July 2003).\\
\underline{Ph D grant} E.U. project: \textit{The impact of Appendicularia in European marine ecosystems (EURAPP)}. Ecology department. University of Oviedo. Spain. (August 1998-October 2000).\\
\underline{Ph D grant} E.U. project: \textit{ Ocean Margin Exchange II - Phase II}. Ecology department. University of Oviedo. Spain. (September 1997-August 1998).\\


\renewcommand{\bibname}{Peer reviewed publications}
\begin{thebibliography}{10}

\bibitem{Alvarez2017}
Alvarez E, Nogueira E, \textbf{L\'{o}pez-Urrutia \'{A}} (2017) In Vivo Single-Cell Fluorescence
  and Size Scaling of Phytoplankton Chlorophyll Content.
\newblock Applied and environmental microbiology 83, \url{Alvarez2017.pdf}.

\bibitem{Gonzalez2017}
Gonzalez P, Alvarez E, Diez J, \textbf{L\'{o}pez-Urrutia \'{A}}, del Coz JJ (2017) Validation
  methods for plankton image classification systems.
\newblock Limnology and oceanography-methods 15:221--237,
  \url{Gonzalez2017.pdf}.

\bibitem{Alvarez2016}
Alvarez E, Mor\'{a}n XAG, \textbf{L\'{o}pez-Urrutia \'{A}}, Nogueira E (2016) Size-dependent
  photoacclimation of the phytoplankton community in temperate shelf waters
  (southern Bay of Biscay).
\newblock Marine ecology progress series 543:73--87, \url{Alvarez2016.pdf}.

\bibitem{Garcia2016}
Garcia FC, Elena Garcia-Martin E, Gonzalez~Taboada F, Sal S, Serret P,
  \textbf{L\'{o}pez-Urrutia \'{A}} (2016) The allometry of the smallest: superlinear scaling of
  microbial metabolic rates in the Atlantic Ocean.
\newblock ISME Journal 10:1029--1036, \url{Garcia2016.pdf}.

\bibitem{Moran2015}
Mor\'{a}n XAG, Alonso-Saez L, Nogueira E, Ducklow HW, Gonzalez N, \textbf{L\'{o}pez-Urrutia \'{A}},
  Diaz-Perez L, Calvo-Diaz A, Arandia-Gorostidi N, Huete-Stauffer TM (2015)
  More, smaller bacteria in response to ocean's warming?
\newblock Proceedings of the Royal Society B-Biological Sciences 282,
  \url{Moran2015.pdf}.

\bibitem{Garcia2015}
Garcia FC, Alonso-Saez L, Mor\'{a}n XAG, \textbf{L\'{o}pez-Urrutia \'{A}} (2015) Seasonality in
  molecular and cytometric diversity of marine bacterioplankton: the
  re-shuffling of bacterial taxa by vertical mixing.
\newblock Environmental microbiology 17:4133--4142, \url{Garcia2015.pdf}.

\bibitem{Sal2015}
Sal S, Alonso-Saez L, Bueno J, Garcia FC, \textbf{L\'{o}pez-Urrutia \'{A}} (2015) Thermal
  adaptation, phylogeny, and the unimodal size scaling of marine phytoplankton
  growth.
\newblock Limnology and oceanography 60:1212--1221, \url{Sal2015.pdf}.

\bibitem{LopezUrrutia2015}
\textbf{L\'{o}pez-Urrutia \'{A}}, Mor\'{a}n XAG (2015) Temperature affects the size-structure of
  phytoplankton communities in the ocean.
\newblock Limnology and oceanography 60:733--738, \url{LopezUrrutia2015.pdf}.

\bibitem{Regaudie2015}
Regaudie-de Gioux A, Sal S, \textbf{L\'{o}pez-Urrutia \'{A}} (2015) Poor correlation between
  phytoplankton community growth rates and nutrient concentration in the sea.
\newblock Biogeosciences 12:1915--1923, \url{Regaudie2015.pdf}.

\bibitem{Vandromme2014}
Vandromme P, Nogueira E, Huret M, \textbf{L\'{o}pez-Urrutia \'{A}}, Gonzalez-Nuevo~Gonzalez G,
  Sourisseau M, Petitgas P (2014) Springtime zooplankton size structure over
  the continental shelf of the Bay of Biscay.
\newblock Ocean Science 10:821--835, \url{Vandromme2014.pdf}.

\bibitem{Garcia2014a}
Garcia FC, \textbf{L\'{o}pez-Urrutia \'{A}}, Mor\'{a}n XAG (2014) Automated clustering of
  heterotrophic bacterioplankton in flow cytometry data.
\newblock Aquatic microbial ecology 72:175--185, \url{Garcia2014a.pdf}.

\bibitem{Alvarez2014a}
Alvarez E, Moyano M, \textbf{L\'{o}pez-Urrutia \'{A}}, Nogueira E, Scharek R (2014) Routine
  determination of plankton community composition and size structure: a
  comparison between FlowCAM and light microscopy.
\newblock Journal of plankton research 36:170--184, \url{Alvarez2014a.pdf}.

\bibitem{Gonzalez2013}
Gonzalez P, Alvarez E, Barranquero J, Diez J, Gonzalez-Quiros R, Nogueira E,
  \textbf{L\'{o}pez-Urrutia \'{A}}, Jose~del Coz J (2013) Multiclass Support Vector Machines
  With Example-Dependent Costs Applied to Plankton Biomass Estimation.
\newblock IEEE Transactions on neural networks and learning systems
  24:1901--1905, \url{Gonzalez2013.pdf}.

\bibitem{Garcia2013a}
Garcia-Munoz C, \textbf{L\'{o}pez-Urrutia \'{A}}, Lubian LM, Garcia CM, Hernandez-Leon S (2013)
  A comparison of primary production models in an area of high mesoscale
  variability (South Shetland Islands, Antarctica).
\newblock Journal of sea research 83:30--39, \url{Garcia2013a.pdf}.

\bibitem{Vazquez2013}
Vazquez-Dominguez E, Mor\'{a}n XAG, \textbf{L\'{o}pez-Urrutia \'{A}} (2013) Photoacclimation of
  picophytoplankton in the central Cantabrian Sea.
\newblock Marine ecology progress series 493:43--56, \url{Vazquez2013.pdf}.

\bibitem{Bueno2012}
Bueno J, \textbf{L\'{o}pez-Urrutia \'{A}} (2012) The Offspring-Development-Time/Offspring-Number
  Trade-Off.
\newblock American naturalist 179:E196--E203, \url{Bueno2012.pdf}.

\bibitem{Alvarez2012a}
Alvarez E, \textbf{L\'{o}pez-Urrutia \'{A}}, Nogueira E (2012) Improvement of plankton biovolume
  estimates derived from image-based automatic sampling devices: application to
  FlowCAM.
\newblock Journal of plankton research 34:454--469, \url{Alvarez2012a.pdf}.

\bibitem{Bode2012}
Bode A, Alvarez-Ossorio MT, Miranda A, \textbf{L\'{o}pez-Urrutia \'{A}}, Vald\'{e}s L (2012)
  Comparing copepod time-series in the north of Spain: Spatial autocorrelation
  of community composition.
\newblock Progress in oceanography 97:108--119, \url{Bode2012.pdf}.

\bibitem{Acuna2011b}
Acu\~{n}a JL, \textbf{L\'{o}pez-Urrutia \'{A}}, Colin S (2011) Primitive Ladder-of-Life Thinking Has
  Evolved Response.
\newblock Science 334:1347--1348, \url{Acuna2011b.pdf}.

\bibitem{Acuna2011}
Acuna J, \textbf{L\'{o}pez-Urrutia \'{A}}, Sean C (2011) Faking giants: the evolution of high
  prey clearance rates in jellyfish.
\newblock Science \url{Acuna2011.pdf}.

\bibitem{Sal2011}
Sal S, \textbf{L\'{o}pez-Urrutia \'{A}} (2011) Temperature, nutrients, and the size-scaling of
  phytoplankton growth in the sea.
\newblock Limnology and Oceanography 56:1952--1955, \url{Sal2011.pdf}.

\bibitem{Rodriguez2011}
Rodriguez J, Gonzalez-Pola C, \textbf{L\'{o}pez-Urrutia \'{A}}, Nogueira E (2011) Composition
  and daytime vertical distribution of the ichthyoplankton assemblage in the
  Central Cantabrian Sea shelf, during summer: An Eulerian study.
\newblock Continental Shelf Research 31:1462--1473, \url{Rodriguez2011.pdf}.

\bibitem{Alvarez2011}
Alvarez E, \textbf{L\'{o}pez-Urrutia \'{A}}, Nogueira E, Fraga S (2011) How to effectively
  sample the plankton size spectrum? A case study using FlowCAM.
\newblock Journal of Plankton Research 33:1119--1133, \url{Alvarez11.pdf}.

\bibitem{Moran2009}
Mor\'{a}n XAG, \textbf{L\'{o}pez-Urrutia \'{A}}, Calvo-Diaz A, W LWK (2009) Increasing importance of
  small phytoplankton in a warmer ocean.
\newblock Global Change Biology
  :10.1111/j.1365--2486.2009.01960.x\url{Moran09.pdf}.

\bibitem{Fernandez2009}
Fernandez~de Puelles ML, \textbf{L\'{o}pez-Urrutia \'{A}}, Morillas A, Molinero JC (2009)
  Seasonal variability of copepod abundance in the Balearic region (Western
  Mediterranean) as an indicator of basin scale hydrological changes.
\newblock Hydrobiologia 617:3--16.

\bibitem{Lopez-Urrutia2008}
\textbf{L\'{o}pez-Urrutia \'{A}} (2008) The metabolic theory of ecology and algal bloom
  formation.
\newblock Limnology And Oceanography 53(5):2046--2047,
  \url{LopezUrrutia08.pdf}.

\bibitem{Benfield2007}
Benfield MC, Grosjean P, Culverhouse PF, Irigoien X, Sieracki ME, Lopez-Urrutia
  A, Dam HG, Hu Q, Davis CS, Hansen A, Pilskaln CH, Riseman EM, Schultz H,
  Utgoff PE, Gorsky G (2007) RAPID Research on Automated Plankton
  Identification.
\newblock Oceanography 20:172--187, \url{Benfield07.pdf}.

\bibitem{Valdes2007}
Vald\'{e}s L, \textbf{L\'{o}pez-Urrutia \'{A}}, Cabal J, Alvarez-Ossorio M, Bode A, Miranda A,
  Cabanas M, Huskin I, Anad\'{o}n R, Alvarez-Marques F, Llope M, Rodriguez N (2007)
  A decade of sampling in the Bay of Biscay: What are the zooplankton time
  series telling us?
\newblock Progress In Oceanography 74:98--114, \url{Valdes08.pdf}.

\bibitem{Lopez-Urrutia2007}
\textbf{L\'{o}pez-Urrutia \'{A}}, Mor\'{a}n XAG (2007) Resource limitation of bacterial production
  distorts the temperature dependence of oceanic carbon cycling.
\newblock Ecology 88:817--822, \url{LopezUrrutiaMoran07.pdf}.

\bibitem{Bonnet2007}
Bonnet D, Harris R, \textbf{L\'{o}pez-Urrutia \'{A}}, Halsband-Lenk C, Greve W, Vald\'{e}s L, Hirche
  HJ, Engel M, Alvarez-Ossorio MT, Wiltshire K (2007) Comparative seasonal
  dynamics of \textit{Centropages typicus} at seven coastal monitoring stations
  in the North Sea, English Channel and Bay of Biscay.
\newblock Progress In Oceanography 72:233--248, \url{Bonnet07.pdf}.

\bibitem{Hirst2006}
Hirst A, \textbf{L\'{o}pez-Urrutia \'{A}} (2006) Effects of evolution on egg development time.
\newblock Marine Ecology-Progress Series 326:29--35,
  \url{HirstLopezUrrutia06.pdf}.

\bibitem{Martin2006}
Martin ES, Irigoien X, Harris RP, \textbf{L\'{o}pez-Urrutia \'{A}}, Zubkov MV, Heywood JL (2006)
  Variation in the transfer of energy in marine plankton along a productivity
  gradient in the Atlantic Ocean.
\newblock Limnology And Oceanography 51:2084--2091, \url{SanMartin06.pdf}.

\bibitem{Varela2006}
Varela M, Bode A, Lorenzo J, Alvarez-Ossorio MT, Miranda A, Patrocinio T,
  Anad\'{o}n R, Viesca L, Rodriguez N, Vald\'{e}s L, Cabal J, \textbf{L\'{o}pez-Urrutia \'{A}},
  Garcia-Soto C, Rodriguez M, Alvarez-Salgado XA, Groom S (2006) The effect of
  the "Prestige" oil spill on the plankton of the N-NW Spanish coast.
\newblock Marine Pollution Bulletin 53:272--286, \url{Varela06.pdf}.

\bibitem{Lopez-Urrutia2006}
\textbf{L\'{o}pez-Urrutia \'{A}}, San~Martin E, Harris RP, Irigoien X (2006) Scaling the
  metabolic balance of the oceans.
\newblock Proceedings Of The National Academy Of Sciences Of The United States
  Of America 103:8739--8744, \url{LopezUrrutia06.pdf}.

\bibitem{Bonnet2005}
Bonnet D, Richardson A, Harris R, Hirst A, Beaugrand G, Edwards M, Ceballos S,
  Diekman R, \textbf{L\'{o}pez-Urrutia \'{A}}, Vald\'{e}s L, Carlotti F, Molinero JC, Weikert H,
  Greve W, Lucic D, Albaina A, Yahia ND, Umani SF, Miranda A, dos Santos A,
  Cook K, Robinson S, de~Puelles MLF (2005) An overview of \textit{Calanus
  helgolandicus} ecology in European waters.
\newblock Progress In Oceanography 65:1--53, \url{Bonnet05.pdf}.

\bibitem{Lopez-Urrutia2005}
\textbf{L\'{o}pez-Urrutia \'{A}}, Harris R, Acuna J, Baamstedt U, Fyhn HJ, Flood P, Gasser B,
  Gorsky X G~andIrigoien, Martinussen M (2005) A comparison of appendicularian
  seasonal cycles in four contrasting European coastal environments.
\newblock In: Gorsky G, Youngbluth M, editors, Response of marine ecosystems to
  global change: Ecological impact of appendicularians, Contemporary Publishing
  International, Aditions Scientifiques GB. pp. 255--276.
\newblock \url{LopezUrrutia05.pdf}.

\bibitem{Fernandez2004}
Fernandez D, \textbf{L\'{o}pez-Urrutia \'{A}}, Fernandez A, Acu\~{n}a JL, Harris R (2004) Retention
  efficiency of 0.2 to 6 um particles by the appendicularians
  \textit{Oikopleura dioica} and \textit{Fritillaria borealis}.
\newblock Marine Ecology-Progress Series 266:89--101, \url{Fernandez04.pdf}.

\bibitem{Lopez-Urrutia2004}
\textbf{L\'{o}pez-Urrutia \'{A}}, Harris RP, Smith T (2004) Predation by calanoid copepods on
  the appendicularian \textit{Oikopleura dioica}.
\newblock Limnology And Oceanography 49:303--307, \url{LopezUrrutia04.pdf}.

\bibitem{Lopez-Urrutia2003}
\textbf{L\'{o}pez-Urrutia \'{A}} (2003) Allometry: How reliable is the biological time clock?
\newblock Nature 424:269--270, \url{LopezUrrutia03c.pdf}.

\bibitem{Lopez-Urrutia2003a}
\textbf{L\'{o}pez-Urrutia \'{A}}, Irigoien X, Acu\~{n}a JL, Harris R (2003) In situ feeding
  physiology and grazing impact of the appendicularian community in temperate
  waters.
\newblock Marine Ecology-Progress Series 252:125--141,
  \url{LopezUrrutia03a.pdf}.

\bibitem{Lopez-Urrutia2003b}
\textbf{L\'{o}pez-Urrutia \'{A}}, Acu\~{n}a JL, Irigoien X, Harris R (2003) Food limitation and
  growth in temperate epipelagic appendicularians (Tunicata).
\newblock Marine Ecology-Progress Series 252:143--157,
  \url{LopezUrrutia03b.pdf}.

\bibitem{Zubkov2003}
Zubkov MV, \textbf{L\'{o}pez-Urrutia \'{A}} (2003) Effect of appendicularians and copepods on
  bacterioplankton composition and growth in the English Channel.
\newblock Aquatic Microbial Ecology 32:39--46, \url{ZubkovLopezUrrutia03.pdf}.

\bibitem{Lopez-Urrutia1999}
\textbf{L\'{o}pez-Urrutia \'{A}}, Acu\~{n}a JL (1999) Gut throughput dynamics in the appendicularian
  \textit{Oikopleura dioica}.
\newblock Marine Ecology-Progress Series 191:195--205,
  \url{LopezUrrutiaAcuna99.pdf}.

\end{thebibliography}

\header{Grants and Fellowships}
\underline{MICINN, Spanish Ministry of Science} \textit{The Malaspina Expedition 2010: Global Change and Exploration of Global Ocean Biodiversity}. (2010-) 594.327 (all IEO)\\

\underline{MICINN, Spanish Ministry of Science} \textit{METOCA: Predicting the metabolic balance of the oceans}. (2010-) 108900 euros.\\

\underline{EU Framework 7} \textit{Marine Ecosystem Evolution in a Changing Environment- MEECE}. (2008-) 115000 euros.\\
\underline{FICTY} \textit{Integration of Methods of image Analysis of planktonic Groups with techniques of Artificial Intelligence}. (2009-2010) 89807 euros.\\
\underline{Effect of hydrographical disturbance on the structure of planktonic communities} \textit{Efecto de las perturbaciones meteorológico-hidrográficas en la estructura de la comunidad planctónica- PERPLAN}. (2007-2010) 98000 euros.\\
\underline{Ramon y Cajal grant} \textit{Marine macroecology: the metabolic theory of ecology and oceanic carbon fluxes}. (2006-2007) 15000 euros.\\
\underline{Study of the effects of ``Prestige'' oil spill on zooplankton and icthyoplankton}. (2003-2006) 95000 euros.\\

\header{Organized Workshops/Symposia}
\underline{ ``Size structure of planktonic communities''}. Session chair, Special Session, ASLO Summer meeting in Santiago de Compostela, 2005.\\
\underline{GLOBEC /SPACC workshop} on ``Image analysis to count and identify zooplankton''. San Sebastian, 2005.\\
\underline{PICES/ICES Early Careers Scientist conference} ``New frontiers in marine science''. June, 2007.\\


\header{Working Groups}
\underline{ICES Working Group on Zooplankton Ecology} Member since 2002.\\
\underline{SCOR Working 125} Global comparison of zooplankton time series. Associate member. 2004-2007.\\
\underline{SCOR Working 113}. Automatic Visual Plankton Identification. Full member. 2007-2010.\\

\header{Invited Talks and Seminars}
\underline{University of Oviedo.} Seminars in ecology. ``Alometría de la producción y la respiración oceánicas y su papel en la modulación del  cambio climático. Oviedo 2006.\\
\underline{European Geophysical Union 2007.} Marine macroecology: simple rules for complex systems. Viena.\\
\underline{GLOBEC IMBER symposium.} Marine macroecology: simple rules for complex systems. Valencia 2007\\
\underline{PICES/ICES Symposium}. Effects of Climate Change on the World's Oceans. Temperature rules the oceans biota.Marine macroecology: simple rules for complex systems. May 2008\\
\header{Profesional OutReach}
\underline{Reviewer}: Trends in Ecology and Evolution, Proccedings of the National Academy of Sciences USA, Ecology, Ecology Letters,  Marine Ecology Progress Series, Marine Biology, Journal of Plankton Research, ICES Journal of Marine Science, Limnology and Oceanography, Journal of the Marine Biological Association UK, Global Change Biology, Oikos, Aquatic Microbial Ecology.\\
\underline{Proposal reviewer} for  NERC UK, United States-Israel Binational Science Foundation, Council for the Earth and Life sciences - Netherlands Organisation for Scientific Research.\\
\underline{Ph.D. panel}: Alejandro Isla (University of Oviedo), Eva Lopez (University of Oviedo), Aitor Alabina (University of the Basque Country) and Lucia Zarautz (University of the Basque Country).\\

\underline{Ph.D. lecturer} for the University of Oviedo Ph.D. programme: Evolutionary and biogeochemical aspects of aquatic systems.

\end{document}

